---
title: "disease_data_info"
author: "Lupe Antonio"
date: "11/30/2023"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(readr)
library(janitor)
```

```{r}
#loading disease dataset & cleaning data
disease_solo <- read_csv('data/raw_data/disease_hospital_admin.csv',
                                   col_types = cols(
                                     `date` = col_date(format = "%m/%d/%Y"),
                                     `total_ed_visits` = col_integer(),
                                     `ili_pne_visits` = col_integer(),
                                     `ili_pne_admissions` = col_integer())) %>%
  separate(date, into = c('year', 'month', 'day')) %>%
  mutate(month = case_when(
    month == '01' ~ 'January',
    month == '02' ~ 'February',
    month == '03' ~ 'March',
    month == '04' ~ 'April',
    month == '05' ~ 'May',
    month == '06' ~ 'June',
    month == '07' ~ 'July',
    month == '08' ~ 'August',
    month == '09' ~ 'September',
    month == '10' ~ 'October',
    month == '11' ~ 'November',
    month == '12' ~ 'December')) %>%
  rename(zip_code = mod_zcta)
```

- `extract_date`: using 12/05/2022, this is the last extraction and most up-to-date data

- `date`: date of emergency dept. visit

- `mod_zcta`: modified zip code tabulation area of patient residence

- `total_ed_vists`: count of all emergency dept visits

- `ili_pne_visits`: count of influenza-like illness and/or pneumonia emergency dept visits

- `ili_pne_admissions`: count of influenza-like illness and/or pneumonia emergency dept visits that resulted in hospital admissions


```{r}
#loading zip shapes file
zip_shapes <- read_csv('data/cleaned_shapes/nyc_zip_codes.csv') %>%
  clean_names()
```

