---
title: "Exploratory Analysis and Visualizations"
output: 
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
---

```{r, echo = FALSE, message = FALSE}
library(tidyverse)
library(leaflet)
library(plotly)
options(readr.show_col_types = FALSE)
```

# Air Pollutant EDA

The Air Quality data is split into seasonal and yearly data, which gives us an idea of pollution in all of New York City's neighborhoods, and daily data, which is more precise with regards to time and location where the measurement was done, but is not available for all neighborhoods in New York City.

## Seasonal and Yearly Data

From the histogram and the boxplots below, it appears that PM2.5 concentrations have been steadily decreasing since 2010. The same trend is evident for NO2, but the decrease seems to be much more gradual relative to PM2.5. Finally, Ozone does not appear to be increasing or decreasing over the time period, but further analysis is needed to confirm this.

```{r, message = FALSE, echo = FALSE}
df_histo_pm2.5 = 
  read_csv("data/cleaned_data/uhf_airquality.csv") |> 
  filter(str_detect(time_period, "Annual Average")) |> 
  mutate(year = str_extract(time_period, "\\d{4}")) |> 
  filter(name == "Fine particles (PM 2.5)",
         year != 2009) |> 
  select(year, data_value)

df_histo_NO2 = 
  read_csv("data/cleaned_data/uhf_airquality.csv")  |> 
  filter(str_detect(time_period, "Annual Average")) |> 
  mutate(year = str_extract(time_period, "\\d{4}")) |> 
  filter(name == "Nitrogen dioxide (NO2)",
         year != 2009) |> 
  select(year, data_value) 

df_histo_ozone =
  read_csv("data/cleaned_data/uhf_airquality.csv")  |> 
  filter(name == "Ozone (O3)") |> 
  mutate(year = str_extract(time_period, "\\d{4}")) |> 
  select(year, data_value) |> 
  filter(year != 2009)
```


### Histogram of levels of NO2, O3 and PM2.5 by year

```{r}
all_years = sort(unique(c(as.numeric(df_histo_pm2.5$year),
                           as.numeric(df_histo_NO2$year),
                           as.numeric(df_histo_ozone$year))))

combined_fig = plot_ly()


pollutants_data = list(df_histo_pm2.5, df_histo_NO2, df_histo_ozone)
pollutant_names = c("PM2.5", "NO2", "O3")
pollutant_units = c("PM2.5 Concentration (mcg/m3)", "NO2 Concentration (ppb)", "O3 Concentration (ppb)")


for(i in seq_along(pollutants_data)) {

  for(year_value in all_years) {
    year_data = filter(pollutants_data[[i]], year == as.character(year_value))
    combined_fig = combined_fig |>
      add_histogram(data = year_data, 
                    x = ~data_value, 
                    name = as.character(year_value),
                    marker = list(opacity = 0.6),
                    visible = i == 1) 
  }
}

num_years = length(all_years)
buttons = lapply(seq_along(pollutants_data), function(i) {
  visibility = c(rep(FALSE, num_years * (i - 1)), 
                 rep(TRUE, num_years), 
                 rep(FALSE, num_years * (length(pollutants_data) - i)))
  list(method = "update",
       args = list(list("visible" = visibility),
                   list("title" = paste(pollutant_names[i], "Concentration"),
                        "xaxis.title" = pollutant_units[i])),
       label = pollutant_names[i])
})

combined_fig = combined_fig |>
  layout(title = "Comparison of Pollutant Concentrations",
         barmode = "overlay",
         xaxis = list(title = "Concentration"),
         yaxis = list(title = 'Count of Measurements'),
         updatemenus = list(list(type = "dropdown", buttons = buttons)))

combined_fig
```

### Boxplots of levels of NO2, O3 and PM2.5 by year 

```{r}
add_yearly_boxplots = function(data, years, name_prefix, visible) {
  for(year_value in years) {
    year_data = filter(data, year == as.character(year_value))
    combined_boxplot_fig <<- combined_boxplot_fig |>
      add_trace(data = year_data, 
                y = ~data_value, 
                type = 'box',
                name = paste(name_prefix, year_value),
                visible = visible)
  }
}


combined_boxplot_fig = plot_ly()


add_yearly_boxplots(df_histo_pm2.5, all_years, "PM2.5", TRUE)

add_yearly_boxplots(df_histo_NO2, all_years, "NO2", FALSE)
add_yearly_boxplots(df_histo_ozone, all_years, "O3", FALSE)


buttons = lapply(seq_along(pollutants_data), function(i) {
  visibility = c(rep(i == 1, num_years), rep(i == 2, num_years), rep(i == 3, num_years))
  list(
    method = "update",
    args = list(
      list("visible" = visibility),
      list(
        "title" = paste(pollutant_names[i], "Concentration"),
        "yaxis.title" = pollutant_units[i]
      )
    ),
    label = pollutant_names[i]
  )
})


combined_boxplot_fig = combined_boxplot_fig |>
  layout(
    title = "Comparison of Pollutant Concentrations (Boxplots)",
    yaxis = list(title = pollutant_units[1]), 
    updatemenus = list(list(type = "dropdown", buttons = buttons))
  )

combined_boxplot_fig

```

## Daily Data

In this data, we have 19,503 rows, where each row is a date-measurement site-pollutant combination.

## How do pollutant variables change over time?

After restricting our analysis to roughly 2020 to 2022, we generate visualizations of the distributions of certain pollutants of interest. Throughout the year, we see oscillations in the levels of Ozone and Nitrogen Dioxide throughout the year with the former reaching its highest levels in the summer months and the latter peaking in the winter. Carbon Monoxide also exhibits oscillations, but to a much lower degree and PM2.5 does not appear to exhibit a seasonal pattern.

```{r, out.width = '100%'}
air_qual_df = read_csv("data/cleaned_data/alt_air_data.csv") |> 
  separate(date, into=c("month", "day", "year")) |>
  mutate(day = as.numeric(day),
         month = month.name[as.numeric(month)],
         year = as.numeric(year)) |> 
  filter(!(year == '2020' & month == 'January'),
         !(year == '2020' & month == 'February')) |> 
  rename(zip = zip_code) |> 
  select(year, month, day, zip, pollutant, value) |> 
  group_by(day, month, year, zip, pollutant) |> 
  summarize(value = mean(value))

poll_plot = air_qual_df |> 
  mutate(
    numeric_month = match(month, month.name),
    date = as.Date(paste(year, numeric_month, day, sep = "-"), 
                   format = "%Y-%m-%d"),
    pollutant = case_when(
      pollutant =="CO" ~ "Carbon Monoxide (ppm)",
      pollutant =="PM2.5" ~ "PM2.5 (μg/m3)",
      pollutant =="Ozone" ~ "Ozone (ppm)",
      pollutant =="NO2" ~ "Nitrogen Dioxide (ppb)"),
    across(pollutant, ~factor(., levels=c("Carbon Monoxide (ppm)",
                                          "Ozone (ppm)",
                                          "Nitrogen Dioxide (ppb)",
                                          "PM2.5 (μg/m3)"))),) |> 
  ggplot(aes(x = date, y = value, color = pollutant)) +
  geom_point() +
  geom_smooth(color = "blue") +
  facet_wrap(~ pollutant, ncol = 2, scales="free_y") +
  ggtitle("Daily average of each pollutant by zip") + 
  scale_color_discrete()

poll_plot
```



```{r message = FALSE, echo = FALSE}
daily_air = read_csv(here::here("data/cleaned_data/alt_air_data.csv")) %>%
  mutate(borough = case_when(
    county == "New York" ~ "Manhattan",
    county == "Bronx" ~ "Bronx",
    county == "Kings" ~ "Brooklyn",
    county == "Queens" ~ "Queens",
    county == "Richmond" ~ "Staten Island"
     ),
    pollutant= case_when(
    pollutant == "PM2.5" ~ "PM 2.5",
    pollutant == "Ozone" ~ "O3",
    T ~ pollutant )) %>% 
  filter(pollutant != "CO") %>%
  mutate(
    pollutant = str_c(pollutant, " (", units, ")")
  )
  
```

### Distributions of NO2, O3 and PM2.5 by borough

We see that at the borough-level, the daily measurements are similar for all pollutants. 

```{r}
daily_air %>%
  ggplot() + geom_boxplot(aes(x = borough, y = value, fill = borough)) + 
  labs(x = "Borough", y = "Measurement value", fill = "Borough") +
  theme(axis.text.x  = element_text(angle=15)) +
  facet_wrap(~pollutant, scales = 'free')
```

### Time series plots for select neighborhoods

In these time series plots, each line represents the mean measurement per day for measurements taken in that neighborhood. The neighborhoods follow similar trends: NO2 seems to peak sometime between winter and spring, while O3 peaks in the summers and PM 2.5 peaks in the winter.

```{r}
#this data maps measurement sites to neighborhoods

sites_nghbors = read_csv(here::here("data/cleaned_data/pollutant_per_neighborhood.csv"))
```

```{r message = FALSE, echo = FALSE}
#Where Fresh Meadows is Queens and Willowbrook is in Staten Island
#Hunts Point - Mott Haven is in Bronx

#my neighborhoods on interest:

neighborhoods = c("Washington Heights", "Union Square - Lower East Side",
                  "Willowbrook", "Hunts Point - Mott Haven", "Fresh Meadows")

store_data = vector("list", length = 5)

for (i in 1:length(neighborhoods)) {
  
  n = neighborhoods[i]
  
  n_zips = sites_nghbors %>%
  filter(geo_place_name == n) %>%
  pull(zip_code)
  
  daily_data = daily_air %>%
    filter(zip_code %in% n_zips) %>%
    group_by(date, pollutant) %>%
    summarize(mean_value = mean(value),
              mean_scaled = mean(scaled_value)) %>%
    mutate(neighborhood = n)
  
  store_data[[i]] = daily_data
  
}
```

```{r}
daily_neighborhood_data = bind_rows(store_data[[1]], store_data[[2]]) %>%
  bind_rows(store_data[[3]]) %>%
  bind_rows(store_data[[4]]) %>%
  bind_rows(store_data[[5]]) %>%
  mutate(date = as.Date(date, format = "%m/%d/%Y"))
```

```{r}
daily_neighborhood_data %>%
  filter(pollutant == "NO2 (ppb)") %>%
  ggplot() + geom_line(aes(x = date, y = mean_value,
                            group = neighborhood, color = neighborhood)) +
  scale_color_manual(values = c("red", "blue", 
                                    "green","orange", 
                                    "black")) +
  labs(title = "NO2 measurements per neighborhood", x = "Date", 
       y = "Mean measurement (ppb)", color = "Neighborhood")
```


```{r}
daily_neighborhood_data %>%
  filter(pollutant == "O3 (ppm)") %>%
  ggplot() + geom_line(aes(x = date, y = mean_value,
                            group = neighborhood, color = neighborhood)) +
  scale_color_manual(values = c("red", "blue", 
                                    "green","orange", 
                                    "black")) +
  labs(title = "O3 measurements per neighborhood", x = "Date", 
       y = "Mean measurement (ppm)", color = "Neighborhood")
```

```{r}
daily_neighborhood_data %>%
  filter(pollutant == "PM 2.5 (ug/m3 LC)") %>%
  ggplot() + geom_line(aes(x = date, y = mean_value,
                            group = neighborhood, color = neighborhood)) +
  scale_color_manual(values = c("red", "blue", 
                                    "black",
                                    "green","orange")) +
  labs(title = "PM 2.5 measurements per neighborhood", x = "Date", 
       y = "Mean measurement (ug/m3 LC)", color = "Neighborhood")
```


# Disease EDA

## Boxplots of respiratory related hospitalizations

## How do disease variables change over time?

Let us first visualize how the number of emergency department visits change over time. We can see that daily ER visits seem to peak in winter months except in 2021. This could be due to COVID restrictions.



From the plot below, we can see that the number of emergency department visits related to pneumonia peak during winter seasons. 



```{r, out.width = '100%'}
all_dis_plot_df = read_csv("data/cleaned_data/disease_only.csv") |>
  mutate(
    numeric_month = match(month, month.name),
    date = as.Date(paste(year, numeric_month, day, sep = "-"), format = "%Y-%m-%d")
  ) |> 
  select(-day, -month, -year, -numeric_month, -total_ed_visits) |> 
  pivot_longer(cols = c("count_resp", "count_asth", "ili_pne_visits", "ili_pne_admissions"), 
               names_to = "variable",
               values_to = "count")

dis_plot = all_dis_plot_df |> 
  
  mutate(variable = case_when(
      variable =="count_resp" ~ "Respiratory diseases",
      variable =="count_asth" ~ "Asthma",
      variable =="ili_pne_visits" ~ "Pneumonia (ER visits)",
      variable =="ili_pne_admissions" ~ "Pneumonia (admissions)"),
      across(variable, ~factor(., levels=c("Asthma","Respiratory diseases",
                                           "Pneumonia (ER visits)",
                                           "Pneumonia (admissions)"))))|> 
  ggplot(aes(x = date, y = count, color = variable)) +
  geom_point() +
  geom_smooth(color = "blue") +
  facet_wrap(~ variable, ncol = 2, scales="free_y") +
  ggtitle("Daily counts of disease variables by zip") + 
  scale_color_discrete()

dis_plot
```

