---
title: "Statistical Analysis"
output: 
  html_document:
    toc: true
    toc_float: true
    theme: cosmo
    code_folding: hide
---
```{r, echo=FALSE, message = FALSE, warning = FALSE}
library(tidyverse)
```


## Regression of Fine Particles (PM 2.5) over time

```{r}
plot_data |> 
  filter(name == "Fine particles (PM 2.5)") |> 
  ggplot(aes(x = year, y = mean_dv)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) + 
  labs(title = "PM2.5 Levels Throughout NYC Over Time",
       x = "Date",
       y = "Pollutant Level (mcg/m3)")


filtered_data = plot_data |> 
  filter(name == "Fine particles (PM 2.5)")

model =
  lm(mean_dv ~ year, data = filtered_data)

summary(model)
```

## Probably wont include, not many data points

```{r}
df_air_zips = 
  read_csv("data/cleaned_data/air_quality_zips.csv") |> 
  mutate(zip = as.factor(zip)) |>
  filter(str_detect(time_period, "Annual Average")) |> 
  mutate(year = str_extract(time_period, "\\d{4}")) |> 
  filter(year >= 2016 & year <= 2020)

df_incd = 
  read_csv("data/disease_datasets/incd.csv") |> 
  janitor::clean_names() |> 
  mutate(fips = as.factor(fips))

df_ziptofip =
  read_csv("data/cleaned_shapes/New_York_State_ZIP_Codes-County_FIPS_Cross-Reference.csv") |> janitor::clean_names() |> 
  mutate(zip = zip_code)

air_quality_with_fips =
  merge(df_air_zips, df_ziptofip, by = "zip") |> 
  mutate(fips = county_fips)

df_air_disease_fips =
  merge(air_quality_with_fips, df_incd, by = "fips")

df_air_disease_fips |> 
  group_by(fips) |> 
  summarise(n = n())

model_pm2.5_means =
  df_air_disease_fips |> 
  filter(name == "Fine particles (PM 2.5)") |>
  group_by(zip) |> 
  summarise(mean_dv = mean(data_value))
  
model_pm2.5_data =
  left_join(model_pm2.5_means, df_air_disease_fips, by = "zip")

model_pm2.5 =
  lm(age_adjusted_incidence_rate_rate_note_cases_per_100_000 ~ mean_dv, model_pm2.5_data)

summary(model_pm2.5) 

model_pm2.5_data |> 
  ggplot(aes(x = mean_dv, y = age_adjusted_incidence_rate_rate_note_cases_per_100_000)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) + 
  labs(title = "Incidence of diseases vs mean PM2.5 concentration",
       x = "Mean pollutant level by zip code (mcg/m3)",
       y = "Age adjusted incidence rate")
```

## Asthma related hospitalizations vs. PM2.5 concentration

```{r}
df_air_zips = 
  read_csv("data/cleaned_data/air_quality_zips.csv") |> 
  mutate(zip = as.factor(zip)) |>
  filter(str_detect(time_period, "Annual Average")) |> 
  mutate(year = str_extract(time_period, "\\d{4}")) |> 
  filter(year >= 2016 & year <= 2020)

asthma_nyc = 
  read_csv("data/disease_datasets/asthmanyc_clean.csv") |> 
  janitor::clean_names() |> 
  select(!c(data_note_1, data_note_2, dim1name)) |> 
  filter(dim2value == "All age groups",
         zip != 88888,
         zip != "Citwide") |> 
  mutate(count = as.numeric(count),
         year = date) |> 
  filter(year >= 2016 & year <= 2020)

df_asthma_air = 
  merge(asthma_nyc, df_air_zips, by = c("zip", "year"))

model_asth_pm2.5_data =
  df_asthma_air |> 
  filter(name == "Fine particles (PM 2.5)")

model_asth_pm2.5 =
  lm(count ~ data_value, data = model_asth_pm2.5_data)

summary(model_asth_pm2.5)

model_asth_pm2.5_data |> 
  ggplot(aes(x = data_value, y = count)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE)
```

## Asthma related hospitalizations vs. NO2 concentration

```{r}
model_asth_no2_data =
  df_asthma_air |> 
  filter(name == "Nitrogen dioxide (NO2)")

model_asth_no2 =
  lm(count ~ data_value, data = model_asth_no2_data)

summary(model_asth_no2)

model_asth_no2_data |> 
  ggplot(aes(x = data_value, y = count)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE)
```

## General respiratory hospitalization vs PM2.5

```{r}
respiratory_nyc = 
  read_csv("data/disease_datasets/respiratorynyc_clean.csv") |> 
  janitor::clean_names() |> 
  select(!c(data_note_1, data_note_2, dim1name)) |> 
  filter(dim2value == "All age groups",
         zip != 88888,
         zip != "Citwide") |> 
  mutate(count = as.numeric(count),
         year = date) |> 
  filter(year >= 2016 & year <= 2020)

df_resp_air = 
  merge(respiratory_nyc, df_air_zips, by = c("zip", "year"))

model_resp_pm2.5_data =
  df_resp_air |> 
  filter(name == "Fine particles (PM 2.5)")

model_resp_pm2.5 =
  lm(count ~ data_value, data = model_resp_pm2.5_data)

summary(model_resp_pm2.5)

model_resp_pm2.5_data |> 
  ggplot(aes(x = data_value, y = count)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE)
```

## general respiratory hospitalizations vs. NO2

```{r}
model_resp_NO2_data =
  df_resp_air |> 
  filter(name == "Nitrogen dioxide (NO2)")

model_NO2_resp =
  lm(count ~ data_value, data = model_resp_NO2_data)

summary(model_NO2_resp)

model_resp_NO2_data |> 
  ggplot(aes(x = data_value, y = count)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE)
```

## Ozone and asthma

```{r}
df_ozone =
  read_csv("data/cleaned_data/air_quality_zips.csv") |> 
  filter(name == "Ozone (O3)") |> 
  filter(str_detect(time_period, "2016|2017|2018|2019|2020")) |> 
  mutate(year = str_extract(time_period, "\\d{4}"))

df_asth_ozone = 
  merge(asthma_nyc, df_ozone, by = c("zip", "year"))

model_ozone_asth =
  lm(count ~ data_value, data = df_asth_ozone)

summary(model_ozone_asth)

df_asth_ozone |> 
  ggplot(aes(x = data_value, y = count)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE)
```


## Ozone and respiratory disease (general)

```{r}
df_ozone =
  read_csv("data/cleaned_data/air_quality_zips.csv") |> 
  filter(name == "Ozone (O3)") |> 
  filter(str_detect(time_period, "2016|2017|2018|2019|2020")) |> 
  mutate(year = str_extract(time_period, "\\d{4}"))

df_resp_ozone = 
  merge(respiratory_nyc, df_ozone, by = c("zip", "year"))

model_ozone_resp =
  lm(count ~ data_value, data = df_resp_air)

summary(model_ozone_resp)

df_resp_ozone |> 
  ggplot(aes(x = data_value, y = count)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE)
```

Readings were taken in Summer.










