---
title: "Fixing UHF Air Quality Data"
author: "Gustavo Garcia-Franceschini"
date: "`r Sys.Date()`"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(sf)
library(sfheaders)
```

# Loading UHF Air Quality

```{r}
df_air = read_csv("data/raw_data/air_quality.csv") %>%
  janitor::clean_names() %>%
  select(-message) %>%
  mutate(id = as.character(geo_join_id)) %>%
  filter(geo_type_name == "UHF42")
```

# Loading zip shapes

```{r}
zip_shapes = read_sf(dsn = "raw_shapes/", layer = 'tl_2019_us_zcta510') %>% 
  rename(zip = ZCTA5CE10)
```

# Loading UHF shapes

I am using the `zip_shapes` file so I can project the `air_shapes` from their current projection back to regular long/lat coordinates.

```{r}
air_shapes = read_sf(dsn = "raw_shapes/", layer = "UHF42") %>%
  filter(id != "0") %>%
  st_transform(crs= st_crs(zip_shapes)) %>%
  rename(geo_join_id = id) %>%
  mutate(geo_join_id = as.numeric(geo_join_id))
```

# Now joining the UHF Air Quality Data with the shapefile

```{r}
df_air = df_air %>%
  left_join(air_shapes) %>%
   mutate(x_center = map_dbl(geometry, ~st_point_on_surface(.x)[[1]]),
         y_center = map_dbl(geometry, ~st_point_on_surface(.x)[[2]]),
         intplon10 = x_center, intplat10 = y_center) #I do this so that the 
        #zip_code shape files and UHF air quality shapefiles have the same 
        #variable names for the point in the center of polygons
        #Otherwise, code done with zip_code shapefiles would need to be
        #edited to account for these new variable names
```

# Plotting to make sure it worked

Now each row has a polygon and an `x_center`, `y_center` coordinate: the center of the polygon

```{r}
ggplot(data = df_air) +
    geom_sf(aes(geometry = geometry)) +
  geom_point(aes(x = x_center, y = y_center)) + theme_void()
```

# Saving data

```{r}
df_air %>%
  select(-geometry) %>% 
  write_csv("data/cleaned_data/uhf_airquality.csv")
```


