---
title: "Interactive Maps"
output: 
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
---

Visualizations of the geographic distribution of air pollutants will help inform our later analyses by showing which locations/sites are impacted the most by air pollution. Three air pollutants are mapped on this page: Ozone (O3), fine particles (PM2.5) and Nitrogen Dioxide (NO2). Ozone and Nitrogen dioxide are measured in parts per billion (ppb), while PM2.5 is measured in micrograms per meter cubed ($mcg/m^3$). The points on these maps show where the measurements have been taken from. Feel free to select a year between 2010 and 2021 from each map and uncheck the heatmap overlay if desired. You may also hover over each point to know more about the location, year and annual average.

```{r, echo = FALSE, message = FALSE, warning = FALSE}
library(tidyverse)
library(leaflet)
library(plotly)
library(glue)
library(leaflet.extras)

df_air_geo = 
  read_csv("data/cleaned_data/uhf_airquality.csv") |> 
  mutate(lat = intplat10,
         lon = intplon10) |> 
  select(-intplat10, -intplon10)
```

## 2010-2021: Ozone levels (O3)

Based on the heatmap, it appears that ozone levels are generally at their highest in Lower Manhattan, particularly in Chelsea (Clinton) and Gramercy Park (Murray Hill). High O3 levels are also seen in coastal areas such as the Rockaway Peninsula, where the Gateway National Recreation Area is located. This makes sense since ozone forms at ground level when nitrogen oxides (such as NO2) react with sunlight, a process that accelerates in warmer temperatures and with increased UV-light exposure. 

From our map, there does not appear to be any apparent trend in ozone levels between 2010 and 2021 and levels seem to oscillate without any pattern. This is lower than the standard of 70 ppb by [the U.S. Environmental Protection Agency (EPA)](https://www.epa.gov/sites/default/files/2015-10/documents/overview_of_2015_rule.pdf). A standard of 70 ppb is below the level that causes adverse health effects in the clinical studies, which essentially eliminates exposures that have been shown to cause adverse health effects, protecting 99.5% of children from even single exposures to ozone at 70 ppb. The highest mean from our data is 40.3 ppb, which is much lower than the standard.

```{r, out.width = '100%', message = FALSE, warning = FALSE}
# Extract ozone data
o3_data = df_air_geo |> 
  filter(str_detect(name, "Ozone") & str_detect(name, "\\(")) |> 
  mutate(start_date = as.Date(start_date, format = "%m/%d/%Y"),
         year = year(start_date),
         name = str_remove(name, "(due to).*")) |> 
  select(geo_place_name, name, measure, measure_info, year, data_value, lat, lon) |> 
  group_by(geo_place_name, name, year, lat, lon) |> 
  summarize(zipmean = mean(data_value)) |> 
  filter(year != 2009)

# Create yearly dataframes
yearly_data_frames <- list()
for (k in 2010:2021) {
  year_data = o3_data |> 
    filter(year == k)
  df_name = paste0("o3_", k)
  assign(df_name, year_data)
  yearly_data_frames[[as.character(k)]] = year_data
}

# Create map

palette <- c("#39FF14", "#FFFF33", "#FF7F00", "#E31A1C", "#9F0000")
legend <- c("21.0-25.0", "25.1-30.0", "30.1-35.0", "35.1-40.0", ">40")

m = leaflet() |> addTiles()

year_i = unique(o3_data$year)
for(year in year_i) {
  
  
  yeardata = o3_data[o3_data$year == year,]
  
  
  label_text <- glue(
    "<b>Neighborhood: </b> {yeardata$geo_place_name}<br/>",
    "<b>Year: </b> {yeardata$year}<br/>",
    "<b>Mean: </b> {yeardata$zipmean} ppb<br/>") |> 
    lapply(htmltools::HTML)
  
  
  m = m |>
    addCircleMarkers(data = yeardata, lng = ~lon, lat = ~lat, radius = .5, opacity = 0.8, 
                     color = ~ifelse(between(zipmean, 21, 25), "#39FF14", 
                                     ifelse(between(zipmean, 25.001, 30), "#FFFF33", 
                                            ifelse(between(zipmean, 30.001, 35),"#FF7F00",
                                                   ifelse(between(zipmean, 35.001, 40), "#E31A1C","#9F0000")))),
                     label = label_text, group = as.character(year))
}

m = m |> 
  addHeatmap(data = yeardata,lng = ~lon, lat = ~lat, 
             intensity = 2, blur = 5, max = 4, radius = 10, 
             group = "Heatmap Overlay") |> 
  addLegend(labels = legend, colors = palette, title = "Ozone levels (parts per billion)") |> 
  addLayersControl(overlayGroups = "Heatmap Overlay",
                   baseGroups = year_i,
                   options = layersControlOptions(collapsed = FALSE)) 

m
```


## 2010-2021: Fine Particles PM2.5

From our map, we see that some of the highest levels of PM2.5 are seen in Lower Manhattan. This makes sense, as we would anticipate that traffic congestion would lead to rising vehicle emissions and subsequently higher concentrations of PM2.5. Unlike O3, however, PM2.5 concentrations generally look to decrease from 2010-2021, especially in neighborhoods that are away from the city. This reflects similar trends seen in previous EDA performed on the data. PM2.5 seems to be the lowest in 2020, which could be due to lower congestion levels during COVID. That being said, there is some year-to-year variability and some clusters of zip codes exhibit high levels as late as 2021. 

Following the [standards by the U.S. Environmental Protection Agency (EPA)](https://www.epa.gov/sites/default/files/2016-04/documents/2012_aqi_factsheet.pdf), the primary annual PM2.5 standards is set to 12 micrograms per cubic meter (µg/m3). Exposure to fine particle pollution can cause premature death and harmful cardiovascular effects such as heart attacks and strokes, and is linked to a variety of other significant health problems. From our maps, we do observe areas with PM2.5 levels higher than this standard, especially in neighborhoods in Lower Manhattan. 

```{r, out.width = '100%', message = FALSE, warning = FALSE}
# Extract pm2.5 data
pm25_data = df_air_geo |> 
  filter(str_detect(name, "Fine")) |> 
  mutate(start_date = as.Date(start_date, format = "%m/%d/%Y"),
         year = year(start_date),
         name = str_remove(name, "(due to).*")) |> 
  select(geo_place_name, name, measure, measure_info, year, data_value, lat, lon) |> 
  group_by(geo_place_name, name, year, lat, lon) |> 
  summarize(zipmean = mean(data_value)) |> 
  filter(year != 2008 & year != 2009)

# Create yearly dataframes
yearly_data_frames <- list()
for (k in 2010:2021) {
  year_data = pm25_data |> 
    filter(year == k)
  df_name = paste0("pm25_", k)
  assign(df_name, year_data)
  yearly_data_frames[[as.character(k)]] = year_data
}

# Create map

palette <- c("#39FF14", "#FFFF33", "#FF7F00", "#E31A1C", "#9F0000", "#660000")
legend <- c("5.0-7.0", "7.1-9.0", "9.1-11.0", "11.1-13.0", "13.1-15.0", ">15")

m = leaflet() |> addTiles()

year_i = unique(pm25_data$year)
for(year in year_i) {
  
  
  yeardata = pm25_data[pm25_data$year == year,]
  
  
  label_text <- glue(
    "<b>Neighborhood: </b> {yeardata$geo_place_name}<br/>",
    "<b>Year: </b> {yeardata$year}<br/>",
    "<b>Mean: </b> {round(yeardata$zipmean,3)} mcg/m3<br/>") |> 
    lapply(htmltools::HTML)
  
  
  m = m |>
    addCircleMarkers(data = yeardata, lng = ~lon, lat = ~lat, radius = .5, opacity = 0.8, 
                     color = ~ifelse(between(zipmean, 5, 7), "#39FF14", 
                                     ifelse(between(zipmean, 7.001, 9), "#FFFF33", 
                                            ifelse(between(zipmean, 9.001, 11),"#FF7F00",
                                                   ifelse(between(zipmean, 11.001, 13), "#E31A1C",
                                                          ifelse(between(zipmean, 13.001, 15), "#9F0000", "#660000"))))),
                     label = label_text, group = as.character(year))
}

m = m |> 
  addHeatmap(data = yeardata,lng = ~lon, lat = ~lat, 
             intensity = 2, blur = 5, max = 4, radius = 10, 
             group = "Heatmap Overlay") |> 
  addLegend(labels = legend, colors = palette, title = "PM2.5 levels (mcg/m3)") |> 
  addLayersControl(overlayGroups = "Heatmap Overlay",
                   baseGroups = year_i,
                   options = layersControlOptions(collapsed = FALSE)) 

m
```


## 2010-2021: Nitrogen dioxide (NO2)

Similar to PM2.5, NO2 levels seem to be at their highest throughout Manhattan, especially in Lower Manhattan. It appears that NO2 levels in boroughs surrounding Manhattan decrease between 2010 and 2021, with South Beach and the Rockaway Peninsula showing lowest NO2 levels throughout the years. The [official level of the annual NO2 standard](https://www.epa.gov/no2-pollution/timeline-nitrogen-dioxide-no2-national-ambient-air-quality-standards-naaqs#:~:text=The%20official%20level%20of%20the,to%20the%201%2Dhour%20standard.&text=The%20form%20of%20the%201,daily%20maximum%20NO2%20concentrations.) is 53 ppb, and our maximum NO2 levels among all neighborhoods and years is 39.8 ppb.

```{r, out.width = '100%', message = FALSE, warning = FALSE}
# Extract NO2 data
no2_data = df_air_geo |> 
  filter(str_detect(name, "Nitrogen")) |> 
  mutate(start_date = as.Date(start_date, format = "%m/%d/%Y"),
         year = year(start_date),
         name = str_remove(name, "(due to).*")) |> 
  select(geo_place_name, name, measure, measure_info, year, data_value, lat, lon) |> 
  group_by(geo_place_name, name, year, lat, lon) |> 
  summarize(zipmean = mean(data_value)) |> 
  filter(year != 2008 & year != 2009)

# Create yearly dataframes
yearly_data_frames <- list()
for (k in 2010:2021) {
  year_data = no2_data |> 
    filter(year == k)
  df_name = paste0("no2_", k)
  assign(df_name, year_data)
  yearly_data_frames[[as.character(k)]] = year_data
}

# Create map

palette <- c("#39FF14", "#FFFF33", "#FF7F00", "#E31A1C", "#9F0000", "#660000")
legend <- c("5.0-10.0", "10.1-15.0", "15.0-20.1", "20.1-25.0", "25.1-30.0", ">30")

m = leaflet() |> addTiles()

year_i = unique(no2_data$year)
for(year in year_i) {
  
  
  yeardata = no2_data[no2_data$year == year,]
  
  
  label_text <- glue(
    "<b>Neighborhood: </b> {yeardata$geo_place_name}<br/>",
    "<b>Year: </b> {yeardata$year}<br/>",
    "<b>Mean: </b> {round(yeardata$zipmean,3)} ppb<br/>") |> 
    lapply(htmltools::HTML)
  
  
  m = m |>
    addCircleMarkers(data = yeardata, lng = ~lon, lat = ~lat, radius = .5, opacity = 0.8, 
                     color = ~ifelse(between(zipmean, 5, 10), "#39FF14", 
                                     ifelse(between(zipmean, 10.001, 15), "#FFFF33", 
                                            ifelse(between(zipmean, 15.001, 20),"#FF7F00",
                                                   ifelse(between(zipmean, 20.001, 25), "#E31A1C",
                                                          ifelse(between(zipmean, 25.001, 30), "#9F0000", "#660000"))))),
                     label = label_text, group = as.character(year))
}

m = m |> 
  addHeatmap(data = yeardata,lng = ~lon, lat = ~lat, 
             intensity = 2, blur = 5, max = 4, radius = 10, 
             group = "Heatmap Overlay") |> 
  addLegend(labels = legend, colors = palette, title = "NO2 levels (parts per billion)") |> 
  addLayersControl(overlayGroups = "Heatmap Overlay",
                   baseGroups = year_i,
                   options = layersControlOptions(collapsed = FALSE)) 

m
```


















