---
title: "prelim_disease_eda"
author: "Lupe Antonio"
date: "12/6/2023"
output: html_document
---

```{r setup, include=FALSE, message = FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(readr)
library(janitor)
library(ggplot2)
```

```{r}
#loading disease data
disease_zip <- read_csv(here::here('data/cleaned_data/disease_zip.csv'))

#loading alt/new air quality data
air_quality <- read_csv(here::here('data/cleaned_data/alt_air_data.csv'))


#loading previously joined data w extra disease data
air_disease <- read_csv('data/cleaned_data/final_airqualdis.csv')


#lots of NAs for most recently joined disease data
```


```{r}
#cleaning second disease data
joined_resp <- read_csv(here::here('data/cleaned_data/joined_respiratory.csv'),
                        col_types = cols(
                          `date` = col_date(format = '%m/%d/%y'))) %>%
  separate(date, into = c('year', 'month', 'day')) %>%
  rename(zip_code = zip) %>%
  filter(!(zip_code == 'Citwide'),
         !(zip_code == '88888'),
         age_group == 'All age groups') %>%
  mutate(month = case_when(
    month == '01' ~ 'January',
    month == '02' ~ 'February',
    month == '03' ~ 'March',
    month == '04' ~ 'April',
    month == '05' ~ 'May',
    month == '06' ~ 'June',
    month == '07' ~ 'July',
    month == '08' ~ 'August',
    month == '09' ~ 'September',
    month == '10' ~ 'October',
    month == '11' ~ 'November',
    month == '12' ~ 'December'),
    
    year = as.integer(year),
    zip_code = as.double(zip_code))
```


```{r}
#joining 2nd option (full_join)
disease_data <- full_join(disease_zip, joined_resp, by = c('zip_code', 'year', 'month', 'day')) %>%
  #deleting unness. columns
  select(-extract_date, -age_group) %>%
  #rearranging columns
  select(year, month, day, zip_code, borough, neighborhood, total_ed_visits, ili_pne_visits, ili_pne_admissions, count_resp, count_asth)
```


Notes:

Zipcodes in second disease that were not included in merging bc no data for it in first disease

- 10271 (zero counts throughout all years) (299 rows)

- 10278 (zero counts throughout) (879 rows)

- 10279 (zero counts throughout) (124 rows)

- 10464 (some dates missing)

(Ended up keeping full joined data to preserve data, even though the majority that was in second but not in first had zeros for data. Additionally, keeping the NAs because don't want to alter to zeros because that may not be accurate data.)



### Disease EDA

```{r}
disease_eda <- disease_data %>%
  pivot_longer(cols = 'total_ed_visits':'count_asth',
               names_to = 'illness_counts',
               values_to = 'count')
```

```{r}
#distribution of illness counts
disease_eda %>%
  ggplot(aes(x = count, fill = illness_counts)) +
  geom_histogram() +
  facet_wrap(~illness_counts, scales = 'free')
```

```{r}
#distribution of illness counts by neighborhood
disease_eda %>%
  ggplot(aes(x = borough, y = count, fill = borough)) +
  geom_boxplot() +
  facet_wrap(~illness_counts, scales = 'free') +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```


```{r}

```





