---
title: "prelim_disease_eda"
author: "Lupe Antonio"
date: "12/6/2023"
output: html_document
---

```{r setup, include=FALSE, message = FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(readr)
library(janitor)
library(ggplot2)
```

```{r, message= FALSE}
#loading disease data
disease_zip <- read_csv(here::here('data/cleaned_data/disease_zip.csv'))

#loading previously joined data w extra disease data
air_disease <- read_csv(here::here('data/cleaned_data/final_airqualdis.csv'))

#loading zip code data
zip_shapes <- read_csv(here::here('data/cleaned_shapes/nyc_zip_codes.csv')) %>%
  clean_names()

#loading new air neigh data
air_neighb <- read_csv(here::here('data/cleaned_data/air_neighborhood.csv'))


#lots of NAs for most recently joined disease data
```


```{r}
#cleaning second disease data
joined_resp <- read_csv(here::here('data/raw_data/joined_respiratory.csv'),
                        col_types = cols(
                          `date` = col_date(format = '%m/%d/%y'))) %>%
  separate(date, into = c('year', 'month', 'day')) %>%
  rename(zip_code = zip) %>%
  filter(!(zip_code == 'Citwide'),
         !(zip_code == '88888'),
         age_group == 'All age groups') %>%
  mutate(month = case_when(
    month == '01' ~ 'January',
    month == '02' ~ 'February',
    month == '03' ~ 'March',
    month == '04' ~ 'April',
    month == '05' ~ 'May',
    month == '06' ~ 'June',
    month == '07' ~ 'July',
    month == '08' ~ 'August',
    month == '09' ~ 'September',
    month == '10' ~ 'October',
    month == '11' ~ 'November',
    month == '12' ~ 'December'),
    
    year = as.integer(year),
    zip_code = as.double(zip_code))
```


```{r}
#joining 2nd option (full_join)
disease_data <- full_join(disease_zip, joined_resp, by = c('zip_code', 'year', 'month', 'day')) %>%
  #deleting unness. columns
  select(-extract_date, -age_group, -borough, -neighborhood)
```

```{r}
#adding missing zip code
disease_data <- left_join(disease_data, zip_shapes, by = 'zip_code') %>%
  mutate(
    neighborhood = case_when(
      zip_code == '10069' ~ 'Upper West Side',
      zip_code == '11109' ~ 'Northwest Queens',
      zip_code == '10282' ~ 'Lower Manhattan',
      zip_code == '10271' ~ 'Lower Manhattan',
      zip_code == '10278' ~ 'Lower Manhattan',
      zip_code == '10279' ~ 'Lower Manhattan',
      #zip_code == '11003' ~ 10000 (deal w these)
      #zip_code == '11040'
      TRUE ~ neighborhood),
  
    borough = case_when(
      zip_code == '10069' ~ 'Manhattan',
      zip_code == '11109' ~ 'Queens',
      zip_code == '10282' ~ 'Manhattan',
      zip_code == '10271' ~ 'Manhattan',
      zip_code == '10278' ~ 'Manhattan',
      zip_code == '10279' ~ 'Manhattan',
      TRUE ~ borough)) %>%
  
  filter(!(zip_code == '10000'),
         !(zip_code == '11003'),
         !(zip_code == '11040'))
```

code below used for checking zip codes w NA's in borough and neighb.
```{r}
#zip techn doesnt exist within nyc
disease_data %>%
  filter(zip_code == '10000')
```

```{r}
#not apart of queens borough
disease_data %>%
  filter(zip_code == '11003')
```

```{r}
#not apart of queen borough
disease_data %>%
  filter(zip_code == '11040')
```

```{r}
na_borough <- disease_data %>%
  filter(is.na(borough))
#none of these zipcodes appear in zip_shapes (9 zipcodes)
unique(na_borough$zip_code)
#10271 - no data in first disease
#10278 - no data in first disease
#10279 - ''''
#11003 - ''''
#11040 - ''''
```

```{r}
na_borough_disease1 <- disease_zip %>%
  filter(is.na(borough))

unique(na_borough_disease1$zip_code)
```

Notes:

Zipcodes in second disease that were not included in merging bc no data for it in first disease

- 10271 (zero counts throughout all years) (299 rows)

- 10278 (zero counts throughout) (879 rows)

- 10279 (zero counts throughout) (124 rows)

- 10464 (some dates missing)

(Ended up keeping full joined data to preserve data, even though the majority that was in second but not in first had zeros for data. Additionally, keeping the NAs because don't want to alter to zeros because that may not be accurate data.)



### Disease EDA

```{r}
disease_eda <- disease_data %>%
  pivot_longer(cols = 'total_ed_visits':'count_asth',
               names_to = 'illness_counts',
               values_to = 'count') %>%
  mutate(month_num = match(month, month.name),
         date = as.Date(paste(year, month_num, day, sep = '-'),
                        format = '%Y-%m-%d'),
         
         illness_counts = case_when(
           illness_counts == 'total_ed_visits' ~ 'ED visits (overall)',
           illness_counts == 'ili_pne_visits' ~ 'Pneumonia (ER visits)',
           illness_counts == 'ili_pne_admissions' ~ 'Pneumonia (ER admissions)',
           illness_counts == 'count_resp' ~ 'Respiratory diseases',
           illness_counts == 'count_asth' ~ 'Asthma'))
```

```{r}
#distribution of illness counts
disease_eda %>%
  ggplot(aes(x = count, fill = illness_counts)) +
  geom_histogram(bins = 15) + 
  facet_wrap(~illness_counts, scales = 'free')
```

```{r}
#distribution of illness counts by neighborhood
disease_eda %>%
  ggplot(aes(x = borough, y = count, fill = borough)) +
  geom_boxplot() +
  facet_wrap(~illness_counts, scales = 'free') +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```


### Observing Washington Heights and Union Square

```{r, message = FALSE}
#prepping for time series

sites_neighb <- read_csv(here::here("data/cleaned_data/pollutant_per_neighborhood.csv"))

washington_heights <- sites_neighb %>%
  filter(geo_place_name == 'Washington Heights') %>%
  pull(zip_code)

counts_wh <- disease_eda %>%
  filter(zip_code %in% washington_heights) %>%
  group_by(date, illness_counts) %>%
  summarize(total_counts = sum(count))
```

```{r, message = FALSE}
#neighborhoods of interest
neighborhoods = c("Washington Heights", "Union Square - Lower East Side",
                  "Willowbrook", "Hunts Point - Mott Haven", "Fresh Meadows")

store_data = vector("list", length = 5)

for (i in 1:length(neighborhoods)) {
  
  n = neighborhoods[i]
  
  n_zips = sites_neighb %>%
  filter(geo_place_name == n) %>%
  pull(zip_code)
  
  counts_data = disease_eda %>%
    filter(zip_code %in% n_zips) %>%
    group_by(date, illness_counts) %>%
    summarize(total_counts = sum(count)) %>%
    mutate(neighborhood = n)
  
  store_data[[i]] = counts_data}
```

```{r}
#combining data above
daily_ill_counts_neigh <- bind_rows(store_data[[1]], store_data[[2]]) %>%
  bind_rows(store_data[[3]]) %>%
  bind_rows(store_data[[4]]) %>%
  bind_rows(store_data[[5]]) %>%
  mutate(date = as.Date(date, format = '%Y-%m-%d'))
```


```{r}
#time series...
daily_ill_counts_neigh %>%
  filter(illness_counts == 'Asthma') %>%
  ggplot(aes(x = date, y = total_counts,
             group = neighborhood, color = neighborhood)) +
  geom_line()
#lots of zeros
```


```{r}
daily_ill_counts_neigh %>%
  filter(illness_counts == 'Respiratory diseases') %>%
  ggplot(aes(x = date, y = total_counts,
             group = neighborhood, color = neighborhood)) +
  geom_line()
#lots of zeros
```


```{r}
daily_ill_counts_neigh %>%
  filter(illness_counts == 'Pneumonia (ER visits)') %>%
  ggplot(aes(x = date, y = total_counts,
             group = neighborhood, color = neighborhood)) +
  geom_line()

#decide against pneumonia (ER admissions) bc visits would be higher
#more likely to have more zeros in admissions, and EDA showed it. 
```


```{r}
daily_ill_counts_neigh %>%
  filter(illness_counts == 'ED visits (overall)') %>%
  ggplot(aes(x = date, y = total_counts,
             group = neighborhood, color = neighborhood)) +
  geom_line()
```


```{r}
daily_ill_counts_neigh %>%
  filter(illness_counts != 'ED visits (overall)') %>%
  pivot_wider(names_from = illness_counts, 
              values_from = total_counts) %>%
  mutate(totals = 'Asthma' + 'Respiratory diseases') %>%
  ggplot(aes(x = date, y = totals,
             group = neighborhood, color = neighborhood)) +
  geom_line()
```


### Joining Disease w Air Quality

```{r}
#focusing on dates between March 1, 2020 and November 2022
#disease_air <- full_join(daily_ill_counts_neigh, air_neighb, by = c('date', "neighborhood")) %>%
#  filter(date >= '2020-03-01' & date < '2022-12-01')
```

```{r}
disease_air[!complete.cases(disease_air), ]
#union sq - no data from march to sept 2020
#wash heigh - no data for may 12 2020, june 17/18 2020
```

```{r}
daily_ill_counts_neigh = daily_ill_counts_neigh %>%
  pivot_wider(names_from = illness_counts, values_from = total_counts)

air_neighb = air_neighb %>%
  pivot_wider(names_from = pollutant, values_from = mean_value)

disease_air <- full_join(daily_ill_counts_neigh, air_neighb, by = c('date', "neighborhood")) %>%
  filter(date >= '2020-03-01' & date < '2022-12-01') %>%
  mutate(year = as.factor(format(date, format = '%Y')),
         month = month.name[as.numeric(format(date, format = "%m"))],
         day = as.numeric(format(date, format = "%d")),
         #adding seasons
         season = case_when(
           month %in% c('June', 'July', 'August') ~ 'Summer',
           month %in% c('September', 'October', 'November') ~ 'Fall',
           month %in% c('December', 'January', 'February') ~ 'Winter',
           month %in% c('March', 'April', 'May') ~ 'Spring'))
  

#Ozone & PM2.5 - Wash & Willow
#All pollutants - Fresh Meadows & Hunts Pt
#PM2.5 - Union Square
```


### Regression

```{r}
library(statar)
new_disease_air = disease_air %>%
  filter(neighborhood != "Union Square - Lower East Side")%>%
  select(-NO2, -Asthma, -`Respiratory diseases`) %>%
  arrange(date) %>%
  group_by(neighborhood) %>%
  mutate(lagged_ozone = tlag(Ozone, 1, time = date),
         lagged_pm = tlag(PM2.5, 1, time = date))

sum(complete.cases(new_disease_air))
```

```{r}
new_disease_air %>%
  mutate(y = sqrt(ili_pne_visits)) %>%
  ggplot() + geom_histogram(aes(x= y))

#bc of zeroes
```

```{r}
new_disease_air = new_disease_air %>%
  mutate(sqrt_pne_visits = sqrt(ili_pne_visits))
```


```{r}
fit = lm(sqrt_pne_visits ~ lagged_ozone + lagged_pm + year + season + neighborhood, 
         data = new_disease_air)
```

```{r}
new_disease_air %>%
  modelr::add_residuals(fit) %>% 
  modelr::add_predictions(fit) %>%
  ggplot(aes(x = pred, y = resid)) + geom_point()
```


```{r}
fit |> 
  broom::glance()
```

```{r}
fit |> 
  broom::tidy()
```

```{r}
library(modelr)
modelr::crossv_mc(new_disease_air, 100) %>%
  mutate(
    pollutants = map(train, \(df) lm(sqrt_pne_visits ~ lagged_ozone + 
                                       lagged_pm, data = df)),
    
    time =  map(train, \(df) lm(sqrt_pne_visits ~ year + season + 
                                       year*season, data = df)),
    
    neighborhoods = map(train, \(df) lm(sqrt_pne_visits ~ neighborhood, 
                                        data = df))) %>%
  mutate(
    rmse_pollutants = map2_dbl(pollutants, test, \(mod, df) rmse(model = mod, 
                                                        data = df)),
    
    rmse_time = map2_dbl(time, test, \(mod, df) rmse(model = mod, 
                                                       data = df)),
    
    rmse_neighborhoods = map2_dbl(neighborhoods, test, \(mod, df) rmse(model = mod, 
                                                             data = df))) %>%
  select(starts_with("rmse")) %>%
  pivot_longer(
    everything(),
    names_to = "model", 
    values_to = "rmse",
    names_prefix = "rmse_") %>%
  mutate(model = fct_inorder(model)) %>%
  ggplot(aes(x = model, y = rmse)) + geom_violin() + 
  labs(title = "RMSE of models predicting sqrt(pneumonia visits)")
```


```{r}
set.seed(42)
bootstrap_results = 
  new_disease_air %>%
  modelr::bootstrap(n = 1000) %>%
  mutate(
    models = map(strap, \(df) lm(sqrt_pne_visits ~ neighborhood, 
                                        data = df)),
    estimates = map(models, broom::tidy),
    results = map(models, broom::glance)) %>% 
  select(-strap, -models) %>%
  unnest(estimates, results) %>%
  select(term, estimate, r.squared) %>% 
  pivot_wider(names_from = term, values_from = estimate)
```

```{r}
bootstrap_results %>%
pivot_longer(r.squared:neighborhoodWillowbrook,
             names_to = "estimate", 
               values_to = "value") %>%
  group_by(estimate) %>%
  summarize(
    ci_lower = quantile(value, 0.025, na.rm = T), 
    ci_upper = quantile(value, 0.975, na.rm = T),
    mean = mean(value))
```

